---
title: "progress-report-knousk"
author: "Keaton Knous"
date: "2026-05-04"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Step 1: Scrape the Data

```{r}
library(rvest)
library(tidyverse)
library(readr)
library(Lahman)

url <- "https://www.baseball-reference.com/awards/hof_2026.shtml"
html <- read_html(url)
tables <- html_table(html)

data <- tables[[1]]
actual_col_names <- data[1, ]
colnames(data) <- actual_col_names
data <- data[-1, ]
```
I scraped the 2026 Hall of Fame ballot from Baseball Reference, which contains the full BBWAA voting results. I selected the first table from the page, and fixed the column names since they were stored in the first row of the data.

## Step 2: Clean the Data

```{r}
hof_2026 <- data %>%
  select(Name, Votes, `%vote`) %>%
  mutate(
    yearID = 2026,
    Name = gsub("X-", "", Name),   
    Votes = as.numeric(Votes),
    percent_vote = readr::parse_number(`%vote`),
    inducted = ifelse(percent_vote >= 75, "Y", "N")
  )

head(hof_2026)
```
I kept the Name, Votes, and %vote columns as they contained the most relevant information. I removed the "X-" prefix from player names that did not make the hall of fame. I also identified inducted players as anyone who received 75% or more of the vote.

## Step 3: Compare to Lahman

```{r}
lahman_2026 <- HallOfFame %>%
  filter(yearID == 2026)

#Row
nrow(hof_2026)
nrow(lahman_2026)

#Counts
hof_2026 %>% count(inducted)
lahman_2026 %>% count(votedBy, inducted)
```
The scraped Baseball Reference data has 27 rows while the Lahman data for 2026 only has 8 rows. This is because Baseball Reference shows the full BBWAA ballot of all 27 candidates, while the Lahman HallOfFame table for 2026 only records the Contemporary Era committee votes, which is a separate voting body with a much smaller ballot.

## Step 4: Save the CSV

```{r}
write.csv(hof_2026, file = "HallOfFame_2026.csv", row.names = FALSE)
```
